# Findings {.unnumbered}

### Relationships in the online era: An investigation of social media usage patterns by relationship status

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(plotly)
media <- read_csv('../project-jar-of-chocolate/data.csv')
media <- media%>%
  rename(App = as.factor('7. What social media platforms do you commonly use?'))%>%
  rename(Gender='2. Gender')%>%
  rename(Time='8. What is the average time you spend on social media every day?')%>%
  rename(MStat='3. Relationship Status')
```

Link to Video: 

## Goals
  In this proposal, we are hoping to investigate how usage of various social media differs by relationship status in addition to  several other questions. We will first determine what the most used social media app is and how long individuals spend on it. Then, we will investigate how that differs by gender and relationship status. Relationship status is a demographic that has rarely been explored as a variable in social media usage statistics. If relationship status is found to influence social media usage time, this could indicate an area for further research. Might, for example, the increased overall happiness of being in a relationship decrease average social media usage time?  Or, might divorced (or single) status lead to a decrease in overall happiness--resulting in an increased dependency on (and thus usage of) social media apps?  

## Data

Our data was acquired from the data repository site "Kaggle".  This dataset consists of usage patterns collected across eight social media platforms--including: Facebook, Pinterest, Reddit, Youtube, Discord, Instagram, Twitter, Snapchat.  There are 481 observations of 21 variables--including demographic questions. For the sake of simplicity, we renamed many of the variables in the dataset (often shortening them).


## Takeaways

Of the people surveyed, single individuals spent the most time using social media (see Fig.1). Married and divorced individuals spent the least amount of time on social media sites, while individuals in a relationship had usage levels on par with single individuals (see Fig.1). It is important to note, however, that there were very few divorced individuals who were surveyed (only six), so results concerning that demographic likely aren't as generalizable as for individuals who are single, married, or in a relationship.  This might be because married individuals have more responsibilities (household chores, children to take care of, etc.) meaning that they have the least amount of free time to spend on social media platforms.  Single individuals and those in relationships likely have fewer responsibilities and thus more time to spend on social media platforms.  


```{r, fig.height = 6, echo=FALSE}

#| fig-alt: 'This is a series of side by side bar plots with time spent on social media during the day on the X-axis and proportion of use by marital status on the Y-axis. The time ranges between zero and five plus hours on social media. The marriage status categories are single, married, divorced, or in a relationship. Overall, it appears that those who identify as single have the greatest time spent on social media, with divorced people having the least hours spent, and those in a relationship not far behind, according to this vizualization. It is also important to note that there is only 6 divorced people in this dataset so their social media usage is probably not representative.'


#Time used by marital status 
fig1 <- media%>%
  filter(Gender %in% c('Female', 'Male'))%>%
  mutate(Time=factor(Time))%>%
  mutate(Time=fct_relevel(Time, c('Less than an Hour', 'Between 1 and 2 hours', 'Between 2 and 3 hours', 'Between 3 and 4 hours', 'Between 4 and 5 hours', 'More than 5 hours')))%>%
   mutate(Time=fct_recode(Time, '<1' = 'Less than an Hour', '1-2'='Between 1 and 2 hours', '2-3'='Between 2 and 3 hours', '3-4'='Between 3 and 4 hours', '4-5'='Between 4 and 5 hours', '>5' = 'More than 5 hours'))%>%
  group_by(Time) %>%
  count(MStat) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  ggplot(aes(y = Time, x = n,fill = MStat, text = paste0("Time (hours): ", Time, "<br>Marital Status: ", MStat, "<br>Percentage: ", round(Percentage, 1), "%"))) +
  geom_bar(stat = "identity", alpha=0.85, position='fill')+
  labs(y='Time Spent on Social Media in hours', x='Propotion of Use by Marital Status', title='Fig. 1: Time on Social Media by Marital Status', fill = 'Marital Status')+
  theme_minimal()+
  scale_fill_manual(values = c("Divorced" = "#B0476E", "Married" = "#C9732A", 'Single' = '#3C7D3F', 'In a relationship' = '#4679B3')) +
  theme(text = element_text(hjust = 0.5, size=12, family='serif'), plot.title=element_text(hjust = 0.5), panel.grid.major = element_blank())

ggplotly(fig1,tooltip = "text")  %>% config(displayModeBar = FALSE)
```


When considering gender-based splits in usage, we see that men use Discord, Reddit, and Twitter more, while Facebook, Instagram, Snapchat, TikTok, and Pinterest are used more by women (see Fig.3).  Additionally, women spent overall more time on social media platforms than men (see Fig.2).  The differences in usage by platform likely demonstrate a gender-based preference for different platforms and also might indicate something about gender relations on different platforms--i.e. whether there is any sort of sexist culture on a given platform.  The fact that women use social media platforms far more than men is fairly suprising and indicates an area for future research--to further investigate additional contributing factors. 

```{r, fig.width= 10, fig.height = 6, echo=FALSE}

#| fig-alt: 'This is a series of side by side bar plots with time spent on social media during the day on the X-axis and proportion of use by gender on the Y-axis. The time ranges between zero and five plus hours on social media. It appears that overall there is a slightly higher proportion of females on social media with 5 + hours being roughly 65% female usage, and less than an hour being 80% male usage.'

#Time used by gender 
fig2 <- media%>%
  filter(Gender %in% c('Female', 'Male'))%>%
    mutate(Time=factor(Time))%>%
  mutate(Time=fct_relevel(Time, c('Less than an Hour', 'Between 1 and 2 hours', 'Between 2 and 3 hours', 'Between 3 and 4 hours', 'Between 4 and 5 hours', 'More than 5 hours')))%>%
  mutate(Time=fct_recode(Time, '<1' = 'Less than an Hour', '1-2'='Between 1 and 2 hours', '2-3'='Between 2 and 3 hours', '3-4'='Between 3 and 4 hours', '4-5'='Between 4 and 5 hours', '>5' = 'More than 5 hours'))%>%
  group_by(Time) %>%
  count(Gender) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  ggplot(aes(y = Time,x = n, fill = Gender, text = paste0("Time Spent (hours): ", Time,"<br>Gender: ", Gender,"<br>Percentage: ", round(Percentage, 1), "%"))) +
  geom_bar(stat = "identity", position='fill' , alpha=0.85)+
  labs(y='Time spent during the day on Social Media in hours', x='Propotion of Use by Gender', title='Fig. 2: Time on Social Media by Gender')+
  theme_minimal()+
  scale_fill_manual(values = c("Female" = "#B0476E", "Male" = "#C9732A")) +
  theme(text = element_text(hjust = 0.5, size=12, family='serif'), plot.title=element_text(hjust = 0.5), panel.grid.major = element_blank())

ggplotly(fig2 , tooltip = "text")  %>% config(displayModeBar = FALSE)
```
\
\
```{r, fig.width= 10, fig.height= 6, echo=FALSE}

#| fig-alt: 'This is a series of side by side bar plots with social media platforms on the X-axis and proportion of use by gender on the Y-axis. This vizualization was created by using data from people who recorded multiple app usage, instead of just one platform. It appears that overall there is a slightly higher proportion of females that use social media than males across all platforms. The most popular platform used by females is Pinterest, and the most popular platform by men is Discord, according to this vizualization.' 


#Total 
fig3 <- media %>% 
  mutate(ig = str_detect(App, 'Instagram')) %>% 
  mutate(fb=str_detect(App, 'Facebook'))%>%
  mutate(tw =str_detect(App, 'Twitter'))%>%
  mutate(tt=str_detect(App, 'TikTok'))%>%
  mutate(dc =str_detect(App, 'Discord'))%>%
  mutate(yt=str_detect(App, 'YouTube'))%>%
  mutate(sc=str_detect(App, 'Snapchat'))%>%
  mutate(pi=str_detect(App, 'Pinterest'))%>%
  mutate(re=str_detect(App, 'Reddit'))%>%
  group_by(Gender)%>%
  summarize(Instagram=sum(ig), Facebook=sum(fb), Twitter=sum(tw), TikTok=sum(tt), Discord=sum(dc), Snapchat=sum(sc), Pinterest=sum(pi), Reddit=sum(re))%>%
  filter(Gender %in% c('Female', 'Male'))%>%
  pivot_longer(cols = -Gender, names_to = "Platform", values_to = "count")%>%
  mutate(Platform = fct_relevel(Platform, c( 'Reddit', 'Discord', 'Twitter',  'Facebook', 'Instagram', 'Snapchat', 'TikTok','Pinterest')))%>%
  group_by(Platform) %>%
  mutate(Percentage = count / sum(count) * 100) %>%
  ggplot(aes( y = Platform, fill = Gender, x = count,text = paste0("Platform: ", Platform,"<br>Gender: ", Gender,"<br>Percentage: ", round(Percentage, 1), "%") )) +
  geom_bar(stat = "identity", position='fill' , alpha=0.85)+
  labs(y='Social Media Platform', x='Propotion of Use by Gender', title='Fig. 3: Use of Social Media Platform by Gender')+
  theme_minimal()+
  scale_fill_manual(values = c("Female" = "#B0476E", "Male" = "#C9732A")) +
  theme(text = element_text(hjust = 0.5, size=12, family='serif'), plot.title=element_text(hjust = 0.5), panel.grid.major = element_blank())

ggplotly(fig3, tooltip = "text")  %>% config(displayModeBar = FALSE)
```


## Big Picture
Overall, these results indicated reason for further research on this demographic of relationship status given the trends indicating a diffrence in social media usage by relationship status.  For example, now that there is evidence of a relationship, it would be worthwhile revisiting the question we posed at the beginning of this analysis: might there be a decreased reliance on social media as a source of happiness for individuals who are married/in relationships?    Additionally, the results on gender-based differences by social media usage and platform also indicate areas for further research--albeit this is a demographic that has been more researched. There are some limitations in our data.  Namely, we only have 481 individuals who participated--thus our sample is likely not as representative as it could have been.  This was demonstrated, for example, in the fact that only 6 divorced individuals were accounted for.  Next, the source of data was "unverifiable", so it is unclear how reliable this dataset is.  Finally, we have a lot of helpful visualizations, but some people may prefer numerical visualizations (i.e. statistics) that would be helpful to include in future data analyses. 
